---
title: "2020 Suncorp Super Netball Analysis"
author: "Emily Partridge"
date: "2023-03-12"
output: 
  html_document:
    code_folding: hide
---

```{r setup, r.options=knitr::opts_chunk$set(warnings=FALSE, message=FALSE, fig.width=12, fig.height=8)}
```

```{r PackagesData}
# Load required packages
library(readr)
library(tidyverse)
library(kableExtra)
library(plotly)
library(ggdark)

# Load in the performance data
RawData <- read_csv("Data/Dataset3_Assessment3.csv")
```

# General Play Turnovers

#### Tibble of team general play turnovers
```{r Create Tibble of GPT}
# Create tibble for general play turnovers
RawData %>%
  filter(Statistic == "generalPlayTurnovers")
```

#### 20 athletes with the most season general play turnovers
```{r NeaterTable, warning=FALSE}
# Create a summary table of performance data
RawData %>% 
  filter(Statistic=="generalPlayTurnovers") %>% 
  group_by(Athlete, Team) %>% 
  summarise(Sum = round(sum(Total), 1)) %>% 
  arrange(desc(Sum)) %>%
  head(20) %>%
  mutate(across(where(is.numeric), round, 1)) %>%
  kbl() %>%
  kable_minimal()
```

### Sum of all general play turnovers for the 2020 Suncorp Netball Season
```{r InteractiveFigure1, warning=FALSE}
# Create a plot for sum of GPT over season for each team
RawData %>% filter(Statistic=="generalPlayTurnovers") %>% group_by(Team) %>% summarise(Sum = round(sum(Total), 1)) %>% 
  arrange(desc(Sum)) %>% ggplot(aes(x = reorder(Team, -Sum), y = Sum)) +
  geom_point(size = 10) + geom_segment(aes(x = Team, y = 0, xend = Team, yend = Sum), linetype = "dashed") +
  geom_text(aes(label = Sum), size = 4, colour = "black", check_overlap = TRUE) +
  scale_y_continuous(expand = c(0,0), limits = c(0, 400), breaks = seq(0,500, by = 50)) +
  labs(x = NULL, y = "Number of General Play Turnovers \n", title = "\n Suncorp Super Netball 2020 \n General Play Turnovers \n") +
  dark_theme_gray()  + theme(plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
        plot.background = element_rect(fill = "grey10"), panel.background = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.background = element_blank(),
        axis.title.y = element_text(size = 12, face = "bold"), axis.ticks.x =  element_line(color = "grey30", size = 0.1),
        axis.line.x =  element_line(color = "grey30", size = 0.1), 
        axis.text.x = element_text(size = 12, face = "bold", angle = 45, vjust = 1, hjust = 1),
        axis.ticks.y =  element_line(color = "grey30", size = 0.1), axis.line.y =  element_line(color = "grey30", size = 0.1),
        axis.text.y = element_text(size = 10, face = "bold"),legend.title = element_blank(), legend.position = "none")
```

# Successful Goals From Zone 1

#### 2020 Season goal scorers
```{r NeaterTable2, warning=FALSE}
# Remove zeros from data frame
RawData_new <- filter_if(RawData, is.numeric, all_vars((.) != 0))

# Create a summary table of individual goals in zone 1 data
RawData_new %>% 
  filter(Statistic=="goal_from_zone1") %>% 
  group_by(Athlete, Team) %>% 
  summarise(Sum = round(sum(Total), 1)) %>%
  arrange(desc(Sum)) %>%
  mutate(across(where(is.numeric), round, 1)) %>%
  kbl() %>%
  kable_minimal()
```

### Sum of Top 10 individual successful shots from zone 1
```{r IntegrativePlot2, warning=FALSE}
#  Plot the top 10 athletes for statistics = goal_from_zone1
RawData %>% filter(Statistic=="goal_from_zone1") %>% group_by(Team) %>% summarise(Sum = round(sum(Total), 1)) %>%  
  arrange(desc(Sum)) %>% head(10) %>% ggplot(aes(x = reorder(Team, -Sum), y = Sum)) +
  geom_point(size = 10) + geom_segment(aes(x = Team, y = 0, xend = Team, yend = Sum), linetype = "dashed") +
  geom_text(aes(label = Sum), size = 4, colour = "black", check_overlap = TRUE) +
  scale_y_continuous(expand = c(0,0), limits = c(0, 900), breaks = seq(0,900, by = 100)) +
  labs(x = NULL, y = "Number of Feeds with Attempt \n", title = "\n Suncorp Super Netball 2020 \n Goals from Zone 1 (Individual Players) \n") +
  dark_theme_gray()  + theme(plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
        plot.background = element_rect(fill = "grey10"), panel.background = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.background = element_blank(),
        axis.title.y = element_text(size = 12, face = "bold"), axis.ticks.x =  element_line(color = "grey30", size = 0.1),
        axis.line.x =  element_line(color = "grey30", size = 0.1), 
        axis.text.x = element_text(size = 12, face = "bold", angle = 45, vjust = 1, hjust = 1),
        axis.ticks.y =  element_line(color = "grey30", size = 0.1), axis.line.y =  element_line(color = "grey30", size = 0.1),
        axis.text.y = element_text(size = 10, face = "bold"),legend.title = element_blank(), legend.position = "none")
``` 
--------
--------


#### Question 4a: What was the first thing you did when you loaded the dataset? Summary statistics? A visualisation? Exploratory data analysis?
First thing I completed after importing the data was to create a tibble of the data for exploratory data analysis by looking at the first and last 10 rows using the head() and tail() functions. I filtered the data to look at the two variables of generalPlayTurnovers and goal_from_zone1 using the filter() function. I then used teh ggplot() function to do a quick exploratory plot to view the ranges of the variables 

#### Question 4b: How did you choose to explore the data? Did you perform any summary statistics to help you further explore the data?
For both data variables explored (goal_from_zone1 and generalPlayTurnovers) I used sum() function to get the totals by athlete or team.

#### Question 4c: What packages did you use to visualise data?
tidyverse to use the ggplot2, dyplr, %>% functions to tidy and present the data
kableExtra package was used for the table visualisation
Plotly for the visual plots
ggdark for the plot aesthetic colours that are also colour blind friendly

#### Question 4d: Did you encounter any issues? Errors? How did you overcome them?
When presenting the data in neat tables I had to figure out how to remove the 0 values from each row. I used the filter_if() function to only include data < 0 in the table of goal_from_zone1 performance data.

I also had an issue with the data not being presented in the plot of goal_from_zone1 but no errors indicating why, so took a little while to realise that it was simply due to the y-axis limits needing to be increased to show the data!

#### Question 4e: Do you have any interpretations/ find anything interesting from that dataset that coaches, athletes or performance staff might be able to use in their practice?
